---
title: "Introduction to state space time series analysis"
author: "Alber Sánchez"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r load_data, include=FALSE}
# load data
data(samples2, package = "eotsfilter")

# subset the samples which have a DETER deforestation date
samples.df <- do.call(rbind, lapply(samples2, function(x){x$sample}))
havedeter <- samples.df$view_date != ""
samples <- samples2[havedeter]

# build a list of data.frames
ts.list <- lapply(samples, function(x){
  mtid <- scidbutil::missingtids(x$time_series$time_id)[1]                      # spot the first missing time ID
  return(x$time_series[x$time_series$time_id < mtid, ]/10000)
})

data.1 <- ts.list[[19]][c("ndvi")]
colnames(data.1) <- c("vi")


data.1 <- ts(data.1$vi, start = 2000 + (3 * 23/365.25), frequency = 23)
```


# The local level model
- The level component is allowed to vary in time
- In a linear model is the intercept
- Disturbances: observation and level. Serially and mutually independent, normally distributed, 0 mean and each has its own variances
- Equations
    - observation or measurement 
    - state equation

```{r}

```


## Deterministic level

- level disturbances are 0
- book uses *diffuse initialization*. It estimates the initial value
- In state space models the unknown parameters include the onbservation and statte duisturbance variances, also known as *hyperparameters*


"Using the diffuse initialisation method, the analysis of the log of the number of UK drivers KSI with the deterministic level model yields the following results:"

```{r}
fit <- lm(data.1~1)
res <- residuals(fit)
```

```{r}
# Coefficients
(coefs <- round(as.numeric(coef(fit)), 8))

# Error variance
(error.var <- round(summary(fit)$sigma^2, 8))
```

```{r}
#par(mfrow=c(1,1))
plot.ts(c(data.1), col = "darkgrey", xlab="", ylab = "log KSI", pch=3, cex=0.5, 
        cex.lab = 0.8, cex.axis = 0.7, xlim = c(0,200))
abline(h = coefs[1], col = "blue", lwd = 2, lty=2)
legend("topright", leg = c("VI", " deterministic level"), 
       cex = 0.6, lty = c(1, 2), col = c("darkgrey","blue"), pch=c(3,NA), 
       bty = "y", horiz = T)
# Figure 2.1: Deterministic level.
```

```{r}
#par(mfrow=c(1,1))
plot(ts(residuals(fit)), ylab = "", xlab = "", xlim = c(0,200), 
     col = "darkgrey", lty = 2)
abline(h=0)
# Figure 2.2: Irregular component for deterministic level model.
```


Tests

```{r}
# Normality test - PASS????
shapiro.test(res)
```
```{r}
#Independence - FAIL???
Box.test(res, lag = 15, type = "Ljung")
```


## Stochastic level




```{r}
fn <- function(params){
  dlm::dlmModPoly(order = 1, dV = exp(params[1]) , dW = exp(params[2]))
}
y <- c(data.1)
fit <- dlm::dlmMLE(y, rep(0,2), fn)
mod <- fn(fit$par)
(obs.error.var <- dlm::V(mod))                        # Observation eq error variance 
(state.error.var <- dlm::W(mod))                      # State eq error variance
(mu.1 <- dlm::dropFirst(dlm::dlmFilter(y,mod)$m)[1])  # MLE of the initial value
res <- residuals(dlm::dlmFilter(y, mod), sd = F)
filtered <- dlm::dlmFilter(y, mod)
smoothed <- dlm::dlmSmooth(filtered)
mu <- dlm::dropFirst(smoothed$s)
mu.1 <- mu[1]
```

Estimates
- Observation eq error variance
- State eq error variance
- MLE of the initial value


```{r}
#par(mfrow=c(1,1))
plot.ts(y, col = "darkgrey", xlab="", ylab = "VI",pch=3, cex=0.5, 
        cex.lab=0.8, cex.axis = 0.7, xlim = c(0,200))
lines(mu, col = "blue", lwd = 2, lty = 2)
#lines(ptw::whit2(y, lambda = 100), col = "red", lwd = 2, lty = 2)
for(i in 1:length(y)){
  if(i%%23 == 0){
    abline(v = i, col = "grey10")
  }
}
legend("topright",leg = c("VI", "stochastic level"), cex = 0.6, 
       lty = c(1, 2), col = c("darkgrey","blue"), pch=c(3,NA), bty = "y", 
       horiz = T)
# Figure 2.3: Stochastic level.
```

```{r}
#par(mfrow=c(1,1))
plot.ts(res, ylab="", xlab="", col = "darkgrey", main = "", cex.main = 0.8)
abline(h=0)
```

Tests

```{r}
# Normality test - FAIL??????????????
shapiro.test(res)

# Independence - FAIL?????????????????
Box.test(res, lag = 15, type = "Ljung")

```



## The local level model ....

```{r}
data.2 <- data.1

fn <- function(params){
  dlm::dlmModPoly(order = 1, dV = exp(params[1]) , dW = exp(params[2]))
}
fit <- dlm::dlmMLE(data.2, rep(0,2),fn)
mod <- fn(fit$par)
(obs.error.var <- dlm::V(mod))
(state.error.var <- dlm::W(mod))
filtered <- dlm::dlmFilter(data.2,mod)
smoothed <- dlm::dlmSmooth(filtered)
#mu <- dlm::dropFirst(smoothed$s)
#mu.1 <- mu[1]
#res <- residuals(filtered, sd = F)
```

